### 2014-10-20 tex/NoSQL_table.tex
tex init

\documentclass{article}
\usepackage{graphicx}
\usepackage{pdflscape}

\begin{document}
Hello
\begin{landscape}
Hello
\end{landscape}
\end{document}