### 2014-10-21 tex/NoSQL_table.tex
Dynamo & BigTable removed

\documentclass{article}
\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[a4paper,margin=.5in,landscape]{geometry}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\begin{document}
%\begin{landscape}
% Table generated by Excel2LaTeX from sheet 'Sheet1'
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
\centering
\begin{tabular}{@{}l| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}|@{}}
\toprule
                                    & \textbf{MongoDB}                                                                                                                                                & \textbf{Cassandra}                                                                                                       & \textbf{HBase}                                                                     & \textbf{CouchBase Server}                                                           & \textbf{Cloudera Impala}            \\ \midrule
\textbf{NoSQL-type}                 & Document-store (BSON)                                                                                                                                           & Wide columnar store                                                                                                      & See BigTable                                                                       & Document-store (JSON)                                                               & SQL on top of Hadoop                \\ \midrule
\textbf{Distributed}                & Master-slave, asynchronous replication                                                                                                                          & Masterless, asynchronous replication                                                                                     & Masterslave or multi-master, asynchronous replication.                            & Multi-master                                                                        & Y, not 100\% symmetric              \\ \midrule
\textbf{Storage}                    & 16MB file limit or GridFS                                                                                                                                       &                                                                                                                          & HDFS                                                                               &                                                                                     & HDFS or HBase                       \\ \midrule
\textbf{Indexing}                   & primary \& secondary (single field, compound, multikey, geospatial, text, hashed){[}2{]}                                                                        & Y, primary \& secondary                                                                                                  & Y, primary \& secondary                                                            & Y, primary \& secondary                                                             & Y, primary \& secondary{[}3{]}      \\ \midrule
\textbf{Query optimization}         & Query optimizer + Shard-keys to speed up distributed queries{[}4{]}                                                                                             & Bloom filter{[}5{]} + No JOIN{[}6{]}                                                                                     & Bloom filter{[}7{]}                                                                & memcached                                                                           &                                     \\ \midrule
\textbf{Partitioning}               & Range partitioning based on shard key                                                                                                                           & Consistent hashing and range partitioning                                                                                & Range-partitioning                                                                 & Hashing function determines bucket.                                                 & By default not, but possible{[}8{]} \\ \midrule
\textbf{Consistency}                & Configurable                                                                                                                                                    & Configurable                                                                                                             & Strong{[}9{]}                                                                      & Within cluster: strong; within multiple clusters: eventual                          &                                     \\ \midrule
\textbf{Transactions / Concurrency} & Single document operations are atomic, otherwise multi-document transactions{[}10{]} Concurrent reads, exclusive writes (lock on DB level). Sharding possible & Normally only eventual consistency, but CQL plugin offers Lightweight transactions{[}11{]} Row-level atomicity{[}12{]} & Single-row transactions (ACID), OCC with MVCC for wider scope transactions{[}13{]} & Application can implement optimistic (using CAS) or pessimistic concurrency control &                                     \\ \midrule
\textbf{Open-source?}               & Y                                                                                                                                                               & Y                                                                                                                        & Y                                                                                  & Y                                                                                   & Y                                   \\ \midrule
\textbf{Written in}                 & C++                                                                                                                                                             & Java                                                                                                                     & Java                                                                               &                                                                                     & C++                                 \\ \bottomrule
\end{tabular}
\end{table}

%\end{landscape}
\end{document}