### 2014-11-13 NoSQL_tabel/NoSQL_table.tex
VoltDB sneuvelversie

\documentclass{article}
\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[a4paper,margin=.5in,landscape]{geometry}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\begin{document}
%\begin{landscape}
% Table generated by Excel2LaTeX from sheet 'Sheet1'
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
\centering
\begin{tabular}{@{}L{2.5cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}|@{}}
\cmidrule(l){2-7}
                                                & \textbf{MongoDB}                                                                                                                                                         & \textbf{Cassandra}                                                                                                                & \textbf{HBase}                                                                              & \textbf{CouchBase Server}                                                                    & \textbf{Cloudera Impala}                     & \textbf{VoltDB}                                                                    \\ \midrule
\multicolumn{1}{|l|}{\textbf{Type}}                & Document-store (BSON)                                                                                                                                           & Wide columnar store                                                                                                      & See BigTable                                                                       & Document-store (JSON)                                                               & SQL on top of Hadoop                & In-memory NewSQL                                                          \\ \midrule
\multicolumn{1}{|l|}{\textbf{Distributed}}               & Master-slave, asynchronous replication                                                                                                                          & Masterless, asynchronous replication                                                                                     & Masterslave or multi-master, asynchronous replication.                            & Multi-master                                                                        & Y, not 100\% symmetric              & Masterless, updates executed on all replicas at the same time.             \\ \midrule
\multicolumn{1}{|l|}{\textbf{Storage}}                   & 16MB file limit or GridFS                                                                                                                                       &                                                                                                                          & HDFS                                                                               &                                                                                     & HDFS or HBase                       &                                                                           \\ \midrule
\multicolumn{1}{|l|}{\textbf{Indexing}}                  & primary \& secondary (single field, compound, multikey, geospatial, text, hashed){[}2{]}                                                                        & Y, primary \& secondary                                                                                                  & Y, primary \& secondary                                                            & Y, primary \& secondary                                                             & Y, primary \& secondary{[}3{]}      & Y, primary \& secondary                                                  \\ \midrule
\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer + Shard-keys to speed up distributed queries{[}4{]}                                                                                             & Bloom filter{[}5{]} + No JOIN{[}6{]}                                                                                     & Bloom filter{[}7{]}                                                                & memcached                                                                           &                                     &                                                                           \\ \midrule
\multicolumn{1}{|l|}{\textbf{Partitioning}}              & Range partitioning based on shard key                                                                                                                           & Consistent hashing                                                                                & Range-partitioning                                                                 & Hashing function determines bucket.                                                 & By default not, but possible{[}8{]} & Consistent hashing                                                        \\ \midrule
\multicolumn{1}{|l|}{\textbf{Consistency}}               & Configurable                                                                                                                                                    & Configurable                                                                                                             & Strong                                                                      & Within cluster: strong; within multiple clusters: eventual                          &                                     & Strong                                                                    \\ \midrule
\multicolumn{1}{|l|}{\textbf{Concurrency control}} & Atomic single document operations, otherwise 2-phase commit{[}10{]} Concurrent reads, exclusive writes (lock on DB level) & CQL offers CAS, row-level atomicity & Single-row transactions (ACID), OCC with MVCC for wider scope operations & Application can implement optimistic (using CAS) or pessimistic concurrency control &                                     & ACID + data access serialized and executed in single-threaded environment
 \\\bottomrule
\end{tabular}
\end{table}

%\end{landscape}
\end{document}