### 2014-12-06 paper_dec/NoSQL_table.tex
Impala components + table in paper (decently\!)

\documentclass{article}
\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[a4paper,margin=.5in,landscape]{geometry}
\usepackage{booktabs}
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\begin{document}
\pagenumbering{gobble}
% Table generated by Excel2LaTeX from sheet 'Sheet1'
% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[ht]
\centering
\begin{tabular}{@{}L{2.5cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}| L{3cm}|@{}}
\cmidrule(l){2-7}
                                         & \textbf{MongoDB}       & \textbf{CouchBase Server}    & \textbf{Cassandra}     & \textbf{HBase}     & \textbf{Cloudera Impala}    & \textbf{VoltDB}  \\ \midrule
\multicolumn{1}{|l|}{\textbf{Type}}      & Document-store (BSON)  & Document-store (JSON)        & Wide columnar store    & Wide columnar store   & SQL on top of Hadoop     & In-memory relational NewSQL          \\ \midrule
\multicolumn{1}{|l|}{\textbf{Querying \& API}} & Proprietary language, dynamic queries with JS, rich API, MapReduce    & N1QL, memcached API, MapReduce          & CQL, rich API, MapReduce            & No query language (Hive via workaround), Java API, MapReduce                                                   & SQL-92, MapReduce               & SQL-92, rich API, MapReduce                                                  \\ \midrule
\multicolumn{1}{|l|}{\textbf{Distributed}}               & Master-slave, asynchronous replication   & Multi-master, asynchronous replication  & Masterless, asynchronous replication     & Master-slave or multi-master, asynchronous replication.                              & Masterless              & Masterless, updates executed on all replicas at the same time.             \\ \midrule
\multicolumn{1}{|l|}{\textbf{Storage}}                   & Standard FS's, 16MB document limit    & Standard Unix \& Windows FS's  & Cassandra File System (HDFS compatible)     & HDFS & HDFS or HBase & \\ \midrule
\multicolumn{1}{|l|}{\textbf{Indexing}}                  & Primary \& secondary on every attribute, B-tree    & Primary \& secondary, B-tree  & Primary \& secondary, LSM-tree                                                                                           & Primary \& secondary, LSM-tree                                                                                                         & Primary \& secondary{[}3{]}      & Primary \& secondary. Hash- \& tree-indexes                             \\ \midrule
\multicolumn{1}{|l|}{\textbf{Query optimization}}        & Query optimizer + Shard-keys to speed up distributed queries         & memcached                                                                                         & Bloom filter                                                                                     & Bloom filter                                                                                                                                      & Partition pruning using partition key &                                                                           \\ \midrule
\multicolumn{1}{|l|}{\textbf{Partitioning}}              & Range partitioning based on shard key         & Hashing function                                                                                                                     & Consistent hashing                                                                                & Range-partitioning                                                                                                               & By default not, but possible & Consistent hashing                                                        \\ \midrule
\multicolumn{1}{|l|}{\textbf{Consistency}}               & Configurable     & Within cluster: strong; within multiple clusters: eventual                                                                                                                                                       & Configurable                                                                                                             & Strong                                                                                        &                                     & Strong                                                                    \\ \midrule
\multicolumn{1}{|l|}{\textbf{Concurrency control}} & Atomic single document operations, otherwise 2-phase commit; Concurrent reads, exclusive writes (lock on DB level) & Application can implement optimistic (using CAS) or pessimistic concurrency control& CQL offers CAS, row-level atomicity & Single-row transactions (ACID possible), OCC with MVCC for wider scope operations  &                                     & ACID + data access serialized and executed in single-threaded environment \\ \midrule
\multicolumn{1}{|l|}{\textbf{Open-source?}}              & Y                                                                                                                                                               & Y                                                                                                                        & Y                                                                                  & Y                                                                                   & Y                                   & Y                                                                        
 \\\bottomrule
\end{tabular}
\end{table}

\end{document}