### 2015-02-23 thesis/thesis.tex
thesis text template dink

\documentclass[master=cws,masteroption=gs]{kulemt}
\setup{title={Master thesis Gemini Cassandra},
		author={Brecht Gossel\'e},
		promotor={Prof. Roel Wuyts},
		assessor={},
		assistant={Prof. Roel Wuyts}}
\setup{filingcard,
		translatedtitle=,
		udc=,
		shortabstract={Here comes a very short abstract, containing no more than 500 words. \LaTeX\ commands can be used here.}}
\begin{document}
\mainmatter
Oi
\end{document}