### 2015-03-16 thesis/thesis.tex

\documentclass[master=cws,masteroption=gs]{kulemt}
\setup{title={Master thesis Gemini Cassandra},
		author={Brecht Gossel\'e},
		promotor={Prof. Roel Wuyts},
		assessor={},
		assistant={Prof. Roel Wuyts}}
\setup{filingcard,
		translatedtitle=,
		udc=,
		shortabstract={Here comes a very short abstract, containing no more than 500 words. \LaTeX\ commands can be used here.}}

\usepackage{url}
\usepackage[pdfusetitle,colorlinks,plainpages=false]{hyperref}
\begin{document}

\begin{preface}
\end{preface}

\tableofcontents

\begin{abstract}
\end{abstract}

\mainmatter

\include{cha1}

\bibliographystyle{abbrv}
\bibliography{biblio}
\end{document}