### 2015-04-14 thesis/thesis.tex
tekst: implementatie combinatie subqueries, query_expressions

\documentclass[master=cws,masteroption=gs]{kulemt}
\setup{title={Schaalbare genoomanalyse met NoSQL- databases},
		author={Brecht Gossel\'e},
		promotor={Prof. Dr. Roel Wuyts},
		assessor={},
		assistant={Prof. Dr. Roel Wuyts}}
\setup{filingcard,
		translatedtitle=,
		udc=,
		shortabstract={Here comes a very short abstract, containing no more than 500 words. \LaTeX\ commands can be used here.}}

\usepackage{url}
\usepackage{graphicx}
\usepackage{booktabs}
\graphicspath{ {afbeeldingen/} }
\usepackage{listings}
\usepackage[pdfusetitle,colorlinks,plainpages=false]{hyperref}

\begin{document}

\begin{preface}
\end{preface}

\tableofcontents

\begin{abstract}
\end{abstract}

\mainmatter

\include{inleiding}
\include{NoSQL_overzicht}
\include{nosql_survey}
\include{gemini_beschrijving}
\include{implementatie}

\bibliographystyle{abbrv}
\bibliography{biblio}
\end{document}